### 2018-03-26 dissertation/1-intro/chapter1.tex
updated chapters 1 and 2

This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.
The solution is based on a previously existing algorithm \citep{vernize2015malicious}, which was limited to centralized networks.
In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
These features are found by tracing analogies to social networks, i.e. identifying how and why two nodes (in the case of VANETs, vehicles) would meet more than once and at a somewhat predictable interval.

@@

\section{Document organization}
The remainder of this dissertation is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.

>>>>

The study of networks is tied to graph theory, since graphs are a useful way to generate models of networks, and therefore many concepts of the two fields overlap.
Mathematical, computational and statistical concepts developed for graphs can be translated to be useful for a variety of different types of networks.
%, like, for example, finding articulation vertices in a graph to locate bandwidth bottlenecks in a computer network.
Therefore, this document utilizes notation from graph theory in order to represent various kinds of networks.

This work introduces TruMan, a trust management solution for vehicular networks. 
TruMan combines solutions to two classic graph theory problems, strongly connected components and graph coloring, in order to provide an efficient approach to identifying malicious nodes in a dynamic and decentralized network.
This is based on a previously existing algorithm, called MaNI \citep{vernize2015malicious}, which is limited to centralized networks.

In order to function in a dynamic and decentralized environment, TruMan takes advantage of features that allow the development of trust relationships between members of the network.
Since there is no unifying agent supervising the network, trust relationships are built over time, as nodes move around the network and meet other nodes.

These features are discovered by tracing analogies with social networks.
There might be groups of nodes which meet each other with a predictable frequency and at predictable intervals, for example vehicles which belong to family members, neighbors or work colleagues.
By considering such features, TruMan allows the formation of strong trust relationships,  which in turn facilitate the discovery of malicious members of the network.

%This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.

%In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
%These features are found by tracing analogies to social networks, i.e. identifying how and why two nodes (in the case of VANETs, vehicles) would meet more than once and at a somewhat predictable interval.

@@

%\section{Document organization}
The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.

### 2018-03-26 dissertation/2-complexnetworks/chapter2.tex
updated chapters 1 and 2

Complex networks can describe many systems which are observed in nature and society through a collection of \textit{vertices} (or \textit{nodes}) and \textit{edges} \citep{newmannetworks}.
They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).
The study of networks is tied to graph theory, since graphs are a useful way to generate models of networks, and therefore many concepts of the two fields overlap.
Mathematical, computational and statistical concepts developed for graphs can be translated to be useful for a variety of different types of networks, like, for example, finding articulation vertices in a graph to locate bandwidth bottlenecks in a computer network.

@@

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs).
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
		VANETs, which are special types of MANETs, are introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as it contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
		Trust in information networks can be observed in some instances, like peer-to-peer networks [cite], although its usages are not relevant for this work.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		An example is the brain, which contains a neural network formed by neurons, cells which enable information processing; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.
		In biological networks, it is difficult to clearly define trust, since nodes may not have any sort of awareness or intelligence (such as cells or proteins).
		Regardless, the study of trust in biological networks is not relevant for this proposal.

@@

Consider an undirected graph $G = (V,E)$, which models one complex network of any kind.
The vertices are the members of the network (computers, humans, etc.) and each edge represents a pair of vertices' ability to exchange data freely.
This graph represents the network's \textit{topology}, that is, the basic structure of the network.
Then, there is also a directed graph $T = (V, O)$, called a \textit{trust graph}.
$T$ contains the same nodes as $G$, although its edges represent the degree of trust (or opinion) each node has towards another.

There are two main ways to describe the edges in $O$: they can be binary, either existing when there is trust or not otherwise, or they can hold a specific trust value within a certain range.
This means the shape of $T$ is not necessarily similar to that of $G$.
For instance, two people can have contact with each other but not maintain a trust relationship, thus altering the layout of the trust graph compared to the network topology.

In the following two sections of this chapter, trust in the contexts of social and technological networks is further explained.
Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which is expanded upon in \autoref{chap:vanets}.

@@

Social networks also have the property of carrying trust from one relationship to another:
information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust \citep{newmannetworks}.
Expanding on that property, there is the concept friends of friends \citep{boissevain1974friends}.
If, for example, nodes $A$ and $B$ have mutual trust and are considered friends, then it is reasonable to assume that some of $A$'s trust for $B$ carries over to other nodes that enjoy mutual trust with $B$.
In other words, a friend of a friend can be considered more trustworthy than the average stranger.
This property is similar, although not identical, to transitivity, since trust is diminished for each extra step an origin node needs to reach a destination, and there is also the possibility that one node distrusts another even if they share a mutual friend.
Naturally, social trust is not commutative ($A$ trusting $B$ does not imply that $B$ trusts $A$).

>>>>

Complex networks can describe many systems which are observed in nature and society through a collection of \textit{vertices} (or \textit{nodes}) and \textit{edges} \citep{newmannetworks}.
They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).

@@

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs).
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
		VANETs, which are special types of MANETs, are introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as it contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
		Trust in information networks can be observed in some instances, like peer-to-peer networks, although its usages are not relevant for this work.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		An example is the brain, which contains a neural network formed by neurons; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.
		In biological networks, it is difficult to clearly define trust, since nodes may not have any sort of awareness or intelligence (such as cells or proteins).
		Regardless, the study of trust in biological networks is not relevant for this proposal.

@@

Consider an undirected graph $G = (V, E_G)$, which models one complex network of any kind.
The vertices are the members of the network (computers, humans, etc.) and each edge represents a pair of vertices' ability to exchange data freely.
This graph represents the network's \textit{topology}, that is, the basic structure of the network.
Then, there is also a directed graph $T = (V, E)$, called a \textit{trust graph}.
$T$ contains the same vertices as $G$, although its edges represent the degree of trust (or opinion) each node has towards another.

There are two main ways to describe the edges in $E$: they can be binary, either existing when there is trust or not otherwise, or they can hold a specific trust value within a certain range.
In some cases, an edge $a\rightarrow b$ with value 0 indicates lack of trust, meaning $a$ has no reason to trust $b$.
In others, it indicates \textit{negative trust}, meaning $a$ has reason to believe $b$ is malicious.

Since $G$ and $T$ represent different types of information, the shape of $T$ is not necessarily similar to that of $G$, even though they share the same vertices.
For instance, two people can have contact with each other (therefore sharing an edge in $G$) but not maintain a trust relationship (therefore not sharing an edge in $T$), thus altering the layout of the trust graph compared to the network topology.

In the following two sections of this chapter, trust is further explained in the contexts of social and technological networks.
Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \autoref{chap:vanets}.

@@

Social networks also have the property of carrying trust from one relationship to another:
information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust \citep{newmannetworks}.
Expanding on that property, there is the concept friends of friends \citep{boissevain1974friends}.
If, for example, nodes $a$ and $b$ have mutual trust and are considered friends, then it is reasonable to assume that some of $a$'s trust for $b$ carries over to other nodes that enjoy mutual trust with $b$.
In other words, a friend of a friend can be considered more trustworthy than the average stranger.
This property is similar, although not identical, to transitivity, since trust is diminished for each extra step an origin node needs to reach a destination, and there is also the possibility that one node distrusts another even if they share a mutual friend.
Naturally, social trust is not commutative ($a$ trusting $b$ does not imply that $b$ trusts $a$).
